---
title: "Facets of Wealth Inequality Visualized"
author: "Superb Bilby"
output: github_document
---

```{r load-packages, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(scales)
```

```{r load-data, echo = FALSE}
bil <- read.csv("data/billionaire/Forbes Annual Report Data - 2020.csv", skip = 1) %>% drop_na()
mob <- read.csv("data/wealth-mobility/2-wealth-structure.csv")
```

## Introduction

An important component of Data Visualization is perspective. When there is an extreme outlier in a data visualization, we do all kinds of data visualization magic tricks to help keep everything in perspective. For example, we log our axes so that everything can fit on a single pane. We also might facet our outliers so that they appear on adjacent panes and are more visible because they are being compared to other outliers. For example, consider the below graph:

```{r echo = FALSE}

example_log <- data.frame(name = c("Jeff Bezos Wealth", "UChicago Endowment", "Med. USA Household Wealth"),
                          wealth = c(132800000000,8600000000, 121700))

example_log %>%
  ggplot(aes(y = name, x = wealth)) +
  geom_col(fill = "purple") +
  scale_x_log10() +
  labs(y = "",
       x = "Wealth (Logged Axis)",
       title = "Logged Axis Example of Wealth (Dollars)") +
  theme_minimal()
  
```

Here, we have a few differing levels of wealth. One shows the U.S. Median Household wealth, which is about $121,700. The other two are the University of Chicago Endowment and Jeff Bezos' wealth, which is about 8.6 billion and 132.8 billion dollars, respectively. Thanks to the logging of the axes, we are able to accurately assess what the exact values of each of the categories are. These methods of creating perspective are valuable and important, and we aim to lean into these methods in this project, however, we also aim to lean into the usefulness of not using these perspective tricks. 

A shortcoming of the above graph is that it does a poor job of highlighting the differences between the three observations. Jeff Bezos' wealth is a little over 15 times the size of UChicago's, but the bars are misleadingly similar in size, as is the nature of logged axes. Perhaps even more shocking in the context of this visualization is that Bezos' wealth is well over 1 million times as large the average American household's wealth, but Jeff Bezos' bar is only about 2.5 times as large. On the other hand, if we were not to incorporate a logged axis and if one were to instead create another bar graph showing a billionaire’s wealth, such as Elon Musk’s wealth, next to the combined wealth of 100 randomly sampled Americans, you would not be able to see the bar of even the combined wealth of all those Americans if it were next to Elon Musk's wealth in the visualization.  It would not matter if that bar was 1 randomly sampled person in America or 1000, it would still be one or two pixels in height, the difference being indiscernible, which is the point of some of the visualizations we aim to create, which is: to express the sheer unimaginable and insurmountable magnitude of extreme wealth.

Global wealth and income inequality is at an all time high, however, the magnitude of this inequality is difficult to fathom. Often, the uber wealthy of society are worshiped for their resilience and innovation, when in reality it might do some good to scorn them for their greed and malice. A resource that helps people more easily digest the extreme wealth disparity may shift people away from the deification of the world’s oligarchs, and shift people toward the rally against such rapacious behavior. We aim to answer the following questions in an attempt to clarify the realities of wealth/income inequality, and deploy our answers to a Shiny application for wider accessibility:
Our high level goal is to create an easily digestible application which conveys the true scale of global wealth and income inequality.  

## Repository Orgnization

- `data`: This folder contains data needed for our analysis and visualization
- `proposal`: This folder contains the rmarkdown and markdown of our proposal
- `analysis`: This folder contains all the analysis and visualization files of our project

## Motivation and Questions

Global wealth and income inequality is at an all time high, however, the magnitude of this inequality is difficult to fathom. Often, the uber wealthy of society are worshiped for their resilience and innovation, when in reality they should be scorned for their greed and malice. A resource that helps people more easily digest the extreme wealth disparity may shift people away from the deification of the world's oligarchs, and shift people toward the rally against such rapacious behavior. We aim to answer the following questions in an attempt to clarify the realities of wealth/income inequality, and deploy our answers to a [Shiny](https://shiny.rstudio.com/) application for wider accessibility: 

### 1) Wealth Inequality
- What can the wealth of a billionaire accomplish? 
- How many of America's poorest would it take to match the wealth of America's richest? 
- What are some ways we can begin to conceptualize how much money the wealthiest humans actually have? 
- Does anybody really need so much wealth? 
- What is the Industry/Geolocation distribution of billionaires

### 2) Wealth Mobility
- Dividing United States income levels to 5 quantiles (0%-20%, 20%-40%, 40%-60, 60%-80%, 80%-100%), can we visualize their mobility patterns in an interactive dashboard. 

## Data

To answer these questions through visualization, we will be relying on the following datasets:  

* Billionaire List
[The Forbes Billionaire's List](https://docs.google.com/spreadsheets/d/1JThOdF95e0NYoea5pCWd5UDXLmIGc1Fwj1aajDSYYhk/edit#gid=685599441). The dataset provides information about billionaire's networths, how old they are, what industry they are in, their gender, and their location. 

* Wealth Mobility Dataset
[The Wealth Mobility Dataset](https://github.com/fpfeffer/WealthMobility/blob/master/Data/2-wealth-structure.csv). The dataset breaks down how wealth is divided up by their original and destined quintiles and race. It also provides intra-quartile and quintile breakdowns, such as information about the wealthiest quintile within the poorest quintile.