---
title: "Facets of Wealth Inequality Visualized"
author: "Superb Bilby"
output: github_document
---

```{r load-packages, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
bil <- read.csv("../data/billionaire/Forbes Annual Report Data - 2020.csv", skip = 1) %>% drop_na()
mob <- read.csv("../data/wealth-mobility/2-wealth-structure.csv")
```

## Goal of this Project
Our high level goal is to create an easily digestible website which conveys the true scale of global wealth and income inequality.  

## Motivation and Questions

An important component of Data Visualization is perspective. When there is an extreme outlier in a data visualization, we do all kinds of data visualization magic tricks to help keep everything in perspective. For example, we log our axes so that everthing can fit on pane. We also might facet our outliers so that they appear on adjacent panes and are more visible because they are being compared to other outliers. These methods of creating perspective are valuable and important, and we aim to lean into these methods in this project, however, we also aim to lean into the usefulness of not using these perspective tricks. For example, if one were to create a bar graph showing a billionaire's wealth, such as Elon Musk's wealth, next to the combined wealth of 100 randomly sampled Americans, you would not be able to see the bar of the combined wealth. It would not matter if that bar was 1 randomly sampled person in America or 1000, it would still be one or two pixels in height, the difference being indiscernible, which is the point of the visualization: to express the sheer unimaginable and insurmountable magnitude of extreme wealth.    

Global wealth and income inequality is at an all time high, however, the magnitude of this inequality is difficult to fathom. Often, the uber wealthy of society are worshiped for their resilience and innovation, when in reality they should be scorned for their greed and malice. A resource that helps people more easily digest the extreme wealth disparity may shift people away from the deification of the world's oligarchs, and shift people toward the rally against such rapacious behavior. We aim to answer the following questions in an attempt to clarify the realities of wealth/income inequality, and deploy our answers to a [Shiny](https://shiny.rstudio.com/) application for wider accessibility: 

### 1) Wealth Inequality
- What can the wealth of a billionaire accomplish?   
- How many of our world's poorest would it take to match the wealth of our world's richest? 
- What are some ways we can begin to conceptualize how much money the wealthiest humans actually have? 
- Does anybody really need so much wealth? 
- What is the Industry/Geolocation distribution of billionaires

### 2) Wealth Mobility
- Dividing income levels to 5 quantiles (0%-20%, 20%-40%, 40%-60, 60%-80%, 80%-100%), can we visualize their mobility patterns in an interactive dashboard. 

## Data

To answer these questions through visualization, we will be relying on the following datasets:  

### 1) Billionaire List
[The Forbes Billionaire's List](https://docs.google.com/spreadsheets/d/1JThOdF95e0NYoea5pCWd5UDXLmIGc1Fwj1aajDSYYhk/edit#gid=685599441). The dataset provides information about billionaire's networths, how old they are, what industry they are in, their gender, and their location.  

**Codebook**

|variable |class     |description |
|:--------|:---------|:-----------|
|Name.and.Rank      |character    | Full name       |
|Wealth..Million    |character    | Wealth in million         |
|Surname            |character    | Surname         |
|World.Rank         |integer      | World rank according to wealth        |
|Age                |integer      | Age        |
|Wealth.Source      |character    | Company name or       |
|Industry           |character    | Industry of the wealth source         |
|Gender             |character    | Gender of the billionaire         |
|State              |character    | If this billionaire is from US, their hometown state       |

**Data**
```{r, echo = FALSE}
kable(head(bil))
```

### 2) Wealth Mobility Dataset
[The Wealth Mobility Dataset](https://github.com/fpfeffer/WealthMobility/blob/master/Data/2-wealth-structure.csv). The dataset breaks down how wealth is divided up by their original and destined quintiles and race. It also provides intra-quartile and quintile breakdowns, such as information about the wealthiest quintile within the poorest quintile.

**Codebook**

|variable |class     |description |
|:--------|:---------|:-----------|
|white     |character   | No if Black, Yes if No       |
|o    |character | Original quintile         |
|d    |character | Destined quintile         |
|w    |dummy   | 0 if this Black, 1 if White      |
|o    |integer | Original quintile from 1 to 5        |
|d    |integer | Destined quintile from 1 to 5        |

**Data**
```{r, echo = FALSE}
kable(head(mob))
```

## Weekly Plan of Attack

Below is the delineation of our plan to get our project complete on time.

<span style="color:red; font-weight: bold"> Micah Corning-Myers' Tasks: 💃</span> <br>
<span style="color:blue; font-weight: bold"> Jinfei Zhu's Tasks: 🕺 </span> <br>
<span style="color:purple; font-weight: bold"> Both Team Member's Tasks: 👯</span> <br>

#### Week 0: Brainstorming & Proposals  (May 9 - May 15)


Proposal: Due Wed, May 11
Peer review: Due Fri, May 13


<ol>
  <li style="color:purple; font-weight: bold"> 👯 Offer up at least 1 preliminary idea for a project </li>
  <li style="color:purple; font-weight: bold">👯 Collect at least one dataset for the project </li>
  <li style="color:blue; font-weight: bold"> 🕺 Add the datasets to the repo and fork the project 1 template </li>
  <li style="color:red; font-weight: bold"> 💃 Edit the template accordingly for project 2 and write up the proposal </li>
  <li style="color:blue; font-weight: bold"> 🕺 Make substantive edits and make proposal presentation worthy </li>
</ol>

#### Week 1: Reviews and Visualizations  (May 16 - May 22)

Revised Proposal: Due Wed, May 18


<ol>
  <li style="color:purple; font-weight: bold">👯 Write peer feedbacks and identify/incorporate relevant peer feedback </li>
  <li style="color:purple; font-weight: bold">👯 Data Cleaning & Tidying </li>
  <li style="color:red; font-weight: bold"> 💃 Complete animated time series visualization </li>
  <li style="color:blue; font-weight: bold"> 🕺 Complete Chloropleth Visualization </li>
  <li style="color:red; font-weight: bold"> 💃 Complete Quintiles Static Visualization </li>
  <li style="color:blue; font-weight: bold"> 🕺 Complete Interactive Visualization (user selects parameter) </li>
</ol>


#### Week 2: Shiny, Write Up, Build Presentation  (May 23 - May 29)

Project Draft for peer review: Due Wed, May 25

<ol>
  <li style="color:red; font-weight: bold"> 💃 Deploy Shiny App </li>
  <li style="color:blue; font-weight: bold"> 🕺 Organize presentation relying on Shiny App  </li>
  <li style="color:purple; font-weight: bold">👯 Complete write up </li>
  <li style="color:purple; font-weight: bold">👯 Practice presentation, give mutual feedback, bug hunt </li>
</ol>

## Visualization plan

### 1) Wealth Inequality
- How wealthy is each billionaire
  - Visualize 1 million with a rice/house (or any emoji), and visualize each billionaire's wealth is equal to how many rices/houses/other emojis
  - Visualize how long it will take an average-income person to become billionaire, from top 10 according to GDP produced per capita.
- See billionaire's distribution interactive dashboard
  - Bar Plot of Industries of billionaire
  - Chloropleth of billionaires from US

### 2) Economic Mobility
- Quintiles Static Visualization with dropdown bar to choose each quintile in an interactive dashboard. 
