---
title: "Facets of Wealth Inequality Visualized"
author: "Superb Bilby"
output: github_document
---

```{r load-packages, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
bil <- read.csv("../data/billionaire/Forbes Annual Report Data - 2020.csv", skip = 1) %>% drop_na()
mob <- read.csv("../data/wealth-mobility/2-wealth-structure.csv")
```

## Goal of this Project
Our high level goal is to create an easily digestible website which conveys the true scale of global wealth and income inequality.  

## Motivation and Questions

Global wealth and income inequality is at an all time high, however, the magnitude of this inequality is difficult to fathom. Often, the uber wealthy of society are worshiped for their resilience and innovation, when in reality they should be scorned for their greed and malice. A resource that helps people more easily digest the extreme wealth disparity may shift people away from the deification of the world's oligarchs, and shift people toward the rally against such rapacious behavior. We aim to answer the following questions in an attempt to clarify the realities of wealth/income inequality, and deploy our answers to a [Shiny](https://shiny.rstudio.com/) application for wider accessibility: 

What can the wealth of a billionaire accomplish? How many of our world's poorest would it take to match the wealth of our world's richest? What are some ways we can begin to conceptualize how much money the wealthiest humans actually have? Does anybody really need so much wealth? 

To answer these questions through visualization, we will be relying on the following datasets:  

1) [The Forbes Billionaire's List](https://docs.google.com/spreadsheets/d/1JThOdF95e0NYoea5pCWd5UDXLmIGc1Fwj1aajDSYYhk/edit#gid=685599441). The dataset provides information about billionaire's networths, how old they are, what industry they are in, their gender, and their location.  

```{r, echo = FALSE}
kable(head(bil))
```

2) [The Wealth Mobility Dataset](https://github.com/fpfeffer/WealthMobility/blob/master/Data/2-wealth-structure.csv). The dataset breaks down how wealth is divided up by quintiles, quartiles, and race. It also provides intra-quartile and quintile breakdowns, such as information about the wealthiest quintile within the poorest quintile.

```{r, echo = FALSE}
kable(head(mob))
```

## Weekly Plan of Attack

Below is the delineation of our plan to get our project complete on time.

<span style="color:red; font-weight: bold"> Micah Corning-Myers' Tasks: 💃</span> <br>
<span style="color:blue; font-weight: bold"> Jinfei Zhu's Tasks: 🕺 </span> <br>
<span style="color:purple; font-weight: bold"> Both Team Member's Tasks: 👯</span> <br>

#### Week 0: Brainstorming & Proposals  

<ol>
  <li style="color:purple; font-weight: bold"> 👯 Offer up at least 1 preliminary idea for a project </li>
  <li style="color:purple; font-weight: bold">👯 Collect at least one dataset for the project </li>
  <li style="color:blue; font-weight: bold"> 🕺 Add the datasets to the repo and fork the project 1 template </li>
  <li style="color:red; font-weight: bold"> 💃 Edit the template accordingly for project 2 and write up the proposal </li>
  <li style="color:blue; font-weight: bold"> 🕺 Make substantive edits and make proposal presentation worthy </li>
</ol>

#### Week 1: Reviews and Visualizations  

<ol>
  <li style="color:purple; font-weight: bold">👯 Write peer feedbacks and identify/incorporate relevant peer feedback </li>
  <li style="color:purple; font-weight: bold">👯 Data Cleaning & Tidying </li>
  <li style="color:red; font-weight: bold"> 💃 Complete animated time series visualization </li>
  <li style="color:blue; font-weight: bold"> 🕺 Complete Chloropleth Visualization </li>
  <li style="color:red; font-weight: bold"> 💃 Complete Quintiles Static Visualization </li>
  <li style="color:blue; font-weight: bold"> 🕺 Complete Interactive Visualization (user selects parameter) </li>
</ol>

#### Week 2: Shiny, Write Up, Build Presentation  
<ol>
  <li style="color:red; font-weight: bold"> 💃 Deploy Shiny App </li>
  <li style="color:blue; font-weight: bold"> 🕺 Organize presentation relying on Shiny App  </li>
  <li style="color:purple; font-weight: bold">👯 Complete write up </li>
  <li style="color:purple; font-weight: bold">👯 Practice presentation, give mutual feedback, bug hunt </li>
</ol>



## Questions

The two questions you want to answer.

## Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
