---
title: "GeoSpatial Analysis of Wealth Inequality Visualized"
author: "Superb Bilby"
output: github_document
---

```{r load-packages}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(colorspace)
library(tigris)
library(fmsb)

# set default theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))
```
## Geospatial Distribution
```{r read-data}
bil <- read.csv("../data/billionaire/Forbes Annual Report Data - 2020.csv", skip = 1) %>% drop_na()
mob <- read.csv("../data/wealth-mobility/2-wealth-structure.csv")
# parse wealth into number
bil <- bil %>%
  mutate(Wealth..Millions. = parse_number(Wealth..Millions.))
head(bil)
```
Note: The Choropleth draws very slowly. I decide to draw a png first and put the plot into the shiny app instead of generate it in real time.
```{r data-wrangling}
# read spatial geometry from tigris package
states_sf <- states() %>%
  shift_geometry()

# group by and count billionaire
bil_by_state <- bil %>%
  select(State)%>%
  filter(State %in% state.name)%>%
  group_by(State) %>%
  summarise(n = n())
head(bil_by_state)

# combine data to 
bil_combined <- states_sf %>%
  filter(NAME %in% state.name) %>%
  left_join(bil_by_state, by = c("NAME" = "State")) %>%
  mutate(rate_cut = cut_number(n, n = 4)) 
head(bil_combined)

# Find the states without Billionnaire
states_no_bil <- bil_combined$NAME[is.na(bil_combined$rate_cut)]
states_no_bil
```

"New Hampshire", "Vermont", "Delaware", "New Mexico", "Alabama", "North Dakota", "Alaska" are the seven states without billionaires...

```{r plot-choropleth}  
bil_choropleth <- ggplot(data = bil_combined) +
  geom_sf(mapping = aes(fill = rate_cut)) +
  scale_fill_discrete_sequential(palette = "viridis",
                                 rev = FALSE,
                                 labels = c("[1,2]", "(2,5]", "(5,10]", "(10, 153]", 0)) +
  # scale_fill_discrete_sequential(palette = "viridis",
  #                               rev = FALSE) +
  labs(
    title = "Number of billionaires by state according to Forbes",
    subtitle = "As of Year 2020", 
    caption = "The Forbes Billionaire's List",
    fill = NULL) +
    ggthemes::theme_map()
bil_choropleth
```


```{r save-plot}
ggsave("choropleth.png")
```

## Industry Distribution
```{r industry-eda}
# TODO: Build a shiny app that choose industry and show top 10 billionaire table in that industry
industry_choices <- bil %>%
  distinct(Industry) %>%
  arrange(Industry) %>%
  pull(Industry)
industry_choices
```
```{r industry plot}
bil_by_industry <- bil %>%
  group_by(Industry)%>%
  summarise(n = n(),
            avg_wealth = mean(Wealth..Millions.))
bil_by_industry
num_by_industry_plot <- bil_by_industry %>%
  arrange(desc(n))%>%
  ggplot(mapping = aes(x = n, y = fct_reorder(Industry, n))) +
  # geom_col(aes(fill = Industry), show.legend = FALSE) +
  scale_fill_brewer(palette="BrBG")+
  geom_col() +
  labs(title = "Number of Billionaires by Industry",
       x = "Number of Billionaires",
       y = "Industry")
num_by_industry_plot
```
## Top n billionarie per industry
```{r top-n-billionaire-per-industry}
# input_industry <- "Technology"
input_industry <- "Telecom"

top_5_data <- bil %>%
  filter(Industry == input_industry) %>%
  arrange(desc(Wealth..Millions.))%>%
  head(5)
top_5_data

top_5_plot <- top_5_data%>%
  ggplot(aes(x=Wealth..Millions., y=fct_reorder(Name.and.Rank, Wealth..Millions.))) +
  geom_col() +
  labs(title = paste("Top 5 Billionaire in", input_industry),
       x = "Wealth in Million",
       y = "Name") +
  scale_x_continuous(labels=scales::dollar_format())
top_5_plot
```


### How Rich They Are
```{r interactive-app}
# Choose a billionaire
name <- "Jeff Bezos"
wealth <- bil %>%
  filter(Name.and.Rank == name)%>%
  select(Wealth..Millions.) %>%
  as.numeric()
wealth

# individual price (in million)
us_household_income <- 0.065
uchicago_master <- 0.16
trip_hawaii <- 0.004
aston_martin <- 0.3
house_in_beverly_hill <- 3
name_school <- 300 # Name Chicago booth


# user input
num_us_household <- 5000
num_uchicago_master <- 1000
num_trip_to_hawaii <- 1000
num_aston_martin <- 1000
num_house_in_beverly_hill <- 100
num_name_school <- 100

# total price
total_household <- us_household_income * num_us_household
total_master <- uchicago_master * num_uchicago_master
total_trip <- trip_hawaii * num_trip_to_hawaii
total_car <- aston_martin * num_aston_martin
total_house <- house_in_beverly_hill * num_house_in_beverly_hill
total_fame <- name_school * num_name_school

money_left <- wealth - sum(total_household,
                            total_master,
                            total_trip,
                            total_car,
                            total_house,
                            total_fame)
spending <- c("Household Spending",
              "UChicago Master",
              "International Trip",
              "Aston Martin",
              "House in Beverly Hill",
              "Name UChicago Booth",
              "Money Left")

value <- c(total_household,
                total_master,
                total_trip,
                total_car,
                total_house,
                total_fame,
                money_left)

df <- data.frame("household" = total_household,
                 "education" = total_master,
                 "travel" = total_trip,
                 "car" = total_car,
                 "housing" = total_house,
                 "fame" = total_fame,
                 "money_left" = money_left)
head(df)

df_longer <- df %>%
            pivot_longer(cols = c("household",
                      "education",
                      "travel",
                      "car",
                      "housing",
                      "fame",
                      "money_left"),
                       names_to = "spending", 
                       values_to = "value")


# df_longer %>%
#   ggplot(aes(y=spending, x=value)) +
#   geom_col()
```
```{r}
data <- rbind(rep(max(value),7) , rep(0,7) , df)
data
# Custom the radarChart !
radarchart(data, axistype=1 , 
 
    #custom polygon
    pcol=rgb(0.2,0.5,0.5,0.9) , pfcol=rgb(0.2,0.5,0.5,0.5) , plwd=4 , 
 
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,max(money_left),round(max(value)/5, digits = -4)), cglwd=0.8,
 
    #custom labels
    vlcex=1.5 
    )
```

```{r name-choices}
name_choices <- bil %>%
  distinct(Name.and.Rank) %>%
  arrange(Name.and.Rank) %>%
  pull(Name.and.Rank)
```



## Reference

Price of Choice:
- Radar Chart: https://r-graph-gallery.com/142-basic-radar-chart.html
- Shiny - Nultiple Numerical value: https://community.rstudio.com/t/creating-multiple-numeric-input-according-to-the-variables-of-an-uploaded-dataset/12293
